---
title: "RWorksheet_MATIAS#3b"
output: pdf_document
date: "2023-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r name, echo = TRUE} 

#1a. Write the codes
respon <- c(1:20)
sex <- c(2,2,1,2,2,2,2,2,2,2,1,2,2,2,2,2,2,2,1,2)
occu <- c(1,3,3,3,1,2,3,1,1,1,3,2,1,3,3,1,3,1,2,1)
personhome <- c(5,7,3,8,5,9,6,7,8,4,7,5,4,7,8,8,3,11,7,6)
sib <- c(6,4,4,1,2,1,5,3,1,2,3,2,5,5,2,1,2,5,3,2)
housetype <- c(1,2,3,1,1,3,3,1,2,3,2,3,2,2,3,3,3,3,3,2)


household_data <- data.frame(
  Respondents = respon,
  Sex = sex,
  FatherOccupation = occu,
  PersonAtHome = personhome,
  SiblingsAtSchool = sib,
  HouseType = housetype
)
household_data

#1b. Describe the data.
str(household_data)
summary(household_data)
# the data frame consists of 20 observations(rows) and 6 variables (columns)
# the variables are:
# respondents - which contains a numeric identifier for each respondent
# sex - represents the gender of the respondent (1 for male, 2 for female)
# father's occupation - indicates the father's occupation (1 for farmer, 2 for driver, 3 for others)
# persons at home - represents the number of people at home
# siblings at school - indicates the number of siblings attending school
# type of house - describes the type of house (1 for wood, 2 for semi-concrete, 3 for concrete)

#1c. Is the mean number of siblings attending is 5?
sib_mean <- mean(household_data$SiblingsAtSchool)
sib_mean
# the mean of the number of siblings at school is 2.95 , not 5. 

#1d. Extract the 1st two rows
first2Rows <- household_data[1:2,]
first2Rows

#1e. Extract 3rd and 5th row with 2nd and 4th column.
thirdand5thRows <- household_data[c(3,5),c(2,4)]
thirdand5thRows


#1e. Select the variable types of houses
types_houses <- household_data$HouseType
types_houses


#1g. Select only all Males respondent that their father occupation was farmer.
farmer <- household_data[household_data$Sex == 1 & household_data$FatherOccupation == 1,]
farmer
# there is no observations

#1h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school.

female_respon <- household_data[household_data$SiblingsAtSchool >= 5,]
female_respon
# there are five observations


#2. Write a R program to create an empty data frame. Using the following codes:
df = data.frame(Ints=integer(),
Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE
)

print("Structure of the empty dataframe:")
print(str(df))

#2a. Describe the results.
# df is an empty data frame created with 0 rows and 5 columns
# the columns has the following data type:
# ints = integer
# doubles = double
# characters = character
# logicals = logical
# factors  = factor (0 levels which means empty)
# serves as a template and can be populated with data



#3. Create a .csv file of this. Save it as HouseholdData.csv

new_respon <- c(1:10)
new_sex <- c("Male", "Female", "Female", "Male", "Male", "Female", "Female", "Male", "Female", "Male")
new_occu <- c(1,2,3,3,1,2,2,3,1,3)
new_personHome <- c(5,7,3,8,6,4,4,2,11,6)
new_sib <- c(2,3,0,5,2,3,1,2,6,2)
new_housetype <- c("Wood", "Congrete", "Congrete", "Wood", "Semi-congrete", "Semi-congrete", "Wood", "Semi-congrete", "Semi-congrete", "Congrete")

HouseholdData <- data.frame(
 Respondents = new_respon,
 Sex = new_sex,
 FatherOccupation = new_occu,
 PersonAtHome = new_personHome,
 SiblingsAtSchool = new_sib,
 HouseType = new_housetype
)
HouseholdData

write.csv(HouseholdData, file = "HouseholdData.csv")


#3a. Import the csv file into the R environment.

import <-  read.csv("HouseholdData.csv")
import

#3b. Convert the Sex into factor using factor() function and change it into integer.[Legend: Male = 1 and Female = 2].

import$Sex <- factor(import$Sex, levels = c("Male", "Female"))
import$Sex <- as.integer(import$Sex)
import$Sex

#3c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood = 1; Congrete = 2; Semi-Congrete = 3].

import$HouseType <- factor(import$HouseType, levels = c("Wood", "Congrete", "Semi-congrete"))
import$HouseType <- as.integer(import$HouseType)
import$HouseType

#3d.On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3.

import$FatherOccupation <- factor(import$FatherOccupation, levels = c(1,2,3), labels = c("Farmer", "Driver", "Others"))
import$FatherOccupation

#3e. Select only all females respondent that has a father whose occupation is driver.

femaledriver <- import[import$Sex == 2 & import$FatherOccupation == "Driver",]
femaledriver

#3f. Select the respondents that have greater than or equal to 5 number of siblings attending school.

greater5 <- import[import$SiblingsAtSchool >= 5,]
greater5


#4. Interpret the graph.
# In comparison to the other sentiments, there were more negative sentiments on July 14. This may be a sign that certain events or themes that day attracted unfavorable opinions.

# On July 15, all opinions grew, with the negative sentiment reaching its maximum level. This may indicate that something occurred on that day to cause unfavorable reactions.

# The negative sentiments remained strong on July 17 and 18, while the neutral and positive feelings remained mostly unchanged.

# Even though all sentiments reached their lowest point on July 20, there were still more negative feelings than any previous day. This can mean that nothing occurred that day.

# On July 21, all sentiments increased, with the negative mood reaching its greatest level. This can also mean that something happened on that day that might have caused others to react negatively.

# We may infer from this data that public sentiment is sensitive to outside causes and that it changes over time.

