---
title: "RWorkSheet_Matias#4b"
output: html_document
date: "2023-11-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.Using the for loop
```{r name, echo = TRUE}

vectorA <- c(1,2,3,4,5)

matrixA <- matrix(0,nrow = 5, ncol =5)

  for (j in 1:5)
  for (k in 1:5)
  {
    matrixA[j,k] <- abs (vectorA[j] - vectorA[k])
  }

matrixA
```

2.Print the string "*" using for() function.
```{r name, echo = TRUE}
for (j in 1:5) {
  cat(paste0("\"", rep("*", j), "\""), "\n")
}
```

3.Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements.
```{r name, echo = TRUE}

input <- as.integer(readline("Enter starting number for Fibonacci Sequence: "))

x <- input
y <- 0


cat("Fibonacci sequence starting from", input, ":\n")

repeat {
 
  nextNum <- x + y
 
  if (nextNum > 500){
    break
  }
  cat(nextNum, " ")
  x <- y
  y <- nextNum
 
}
```

4. What is the R script for importing an excel or a csv file?
```{r name, echo = TRUE}

write.csv(HOUSEHOLD_DATA, file = "HOUSEHOLD_DATA.csv")
import<- read.csv("HOUSEHOLD_DATA.csv")
import

imported<-head(HOUSEHOLD_DATA, 6)
imported
```

4b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female?
```{r name, echo = TRUE}

male <- import[import$Gender == "M",]
male

female <- import[import$Gender == "F",]
female

numMale <- nrow(male)
numMale

numFem <- nrow(female)
numFem
```

4c. Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors.
```{r name, echo = TRUE}
totalMaleFemale <- table(import$Gender)
barplot(totalMaleFemale,
        main = "Number of Males and Females",
        xlab = "Gender",
        ylab = "Count",
        col = c("pink", "skyblue"),
        legend.text = rownames(totalMaleFemale),
        beside = TRUE)
```

5. The monthly income of Dela Cruz family was spent on the following:
```{r name, echo = TRUE}

spend_data <- data.frame(
  Category = c("Food", "Electricity", "Savings", "Miscellaneous"),
  Value = c(60, 10, 5, 25)
)

spend_data$Percentage <- spend_data$Value / sum(spend_data$Value) * 100

colors <- c("cyan", "pink", "blue", "purple")

pie(spend_data$Value,
    labels = paste(spend_data$Category, "(",spend_data$Percentage,"%)"),
    col = colors,
    main = "Monthly Income of Dela Cruz Family was spent")

legend("topright", spend_data$Category, fill = colors)
```

6a. Check for the structure of the data set using the str() function
```{r name, echo = TRUE}
data(iris)
str(iris)
```

6b.Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width.
```{r name, echo = TRUE}
mean_Iris <- c(mean(iris$Sepal.Length), mean(iris$Sepal.Width), mean(iris$Petal.Length), mean(iris$Petal.Width))
mean_Iris

meanSepal_Length <- mean(iris$Sepal.Length)
meanSepal_Width <- mean(iris$Sepal.Width)
meanPetal_Length <- mean(iris$Petal.Length)
meanPetal_Width <- mean(iris$Petal.Width)

mean_Iris <- data.frame(Sepal_Length = meanSepal_Length,
                        Sepal_Width = meanSepal_Width,
                        Petal_Length = meanPetal_Length,
                        Petal_Width = meanPetal_Width)
mean_Iris
```

6c. Create a pie chart for the Species distribution. Add title, legends, and colors.
```{r name, echo = TRUE}
Species <- table(iris$Species)
pie(Species, labels = Species, col = rainbow(length(Species)), main = "Species Distribution")
legend("topright", names(Species), cex = 1.0 , fill = rainbow(length(Species)))
```

6d. Subset the species into setosa, versicolor, and virginica. Write the R scripts and show the last six (6) rows of each species.
```{r name, echo = TRUE}
setosa<- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")

tail(setosa, 6)
tail(versicolor, 6)
tail(virginica, 6)
```

6e.Create a scatterplot of the sepal.length and sepal.width using the different species(setosa, versicolor, virginica). Add a title = “Iris Dataset”, subtitle = “Sepal width and length, labels for the x and y axis, the pch symbol and colors should be based on the species.
```{r name, echo = TRUE}
iris$Species <- as.factor(iris$Species)

# Convert the "Species" column to a factor
iris$Species <- as.factor(iris$Species)

# Create a scatterplot
plot(
  Sepal.Length ~ Sepal.Width,
  data = iris,
  pch = as.integer(iris$Species),
  # Use different pch symbols for each species
  col = as.integer(iris$Species),
  # Use different colors for each species
  xlab = "Sepal Length",
  ylab = "Sepal Width",
  main = "Iris Dataset",
  sub = "Sepal width and length"
)

# Add a legend
legend("topright", legend = levels(iris$Species), col = 1:3, pch = 1:3, cex = 0.6, title = "Species")
```