---
title: "RWorksheet_Matias#4c"
output: html_document
date: "2023-11-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1a. Use the dataset mpg. Show your solutions on how to import a csv file into the environment.
```{r}
mpg <- read.csv(file = "mpg.csv", header = T, sep = ",")
mpg
```
1b.
```{r}
mpg_Var <- names(sapply(mpg, function(x) is.factor(x) || is.character(x)))
mpg_Var
```
1c.
```{r}
ContVarsNnames <- names(sapply(mpg, is.numeric))
ContVarsNnames
```
2.Which manufacturer has the most models in this data set? Which model has the most variations?
```{r}
library(dplyr)

  most_manufacturer <- mpg %>%
  group_by(manufacturer) %>%
  summarize(nummodels = n_distinct(model)) %>%
  arrange(desc(nummodels)) %>%
  head(1)


  most_variations <- mpg %>%
  group_by(model) %>%
  summarize(numvariations = n()) %>%
  arrange(desc(numvariations)) %>%
  head(1)


cat("Manufacturer with the Most Models:", most_manufacturer $manufacturer, "\n")
cat("Model with the Most Variations:", most_variations$model, "\n")

```
2a. Group the manufacturers and find the unique models.
```{r}
library(dplyr)

unique_models <- mpg %>%
  group_by(manufacturer) %>%
  summarize(unique_models = n_distinct(model))

print(unique_models)
```
2b. Graph the result by using plot() and ggplot().
```{r}

```
2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.
```{r}
```
2a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
```
2b. For you, is it useful? If not, how could you modify the data to make it more informative?
```{r}
```
3.Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```


